Stack Safety is a Security Property

### 2021-04-15 tex/main.tex

int g() {
  int z;
  if (z == 5)
    return 10;
  else
    return 20;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  {\small
\begin{verbatim}
1 main:  add $1,%sp ; allocate frame
2        call f
3        mov %rax,-1(%sp) ; store x
4        call g
5        sub $1,%sp ; deallocate frame
6        ret

10 f:    add $1,%sp ; allocate frame
11       mov $5,-1(%sp) ; initialize y
12       mov -1(%sp),%rax ; set return value
13       sub $1,%sp ; deallocate frame
14       ret

20 g:    add $1,%sp ; allocate frame
21       cmp $5,-1(%sp) ; compare 
22       brneq #25      ; branch if not equal
23       mov $10,%rax ; set return value
24       jmp #26
25       mov $20,%rax  ; set return value
26       sub $1,%sp ; deallocate frame
27       ret 
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.65\textwidth}
\begin{center}
\begin{tabular}{l r | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
%    \MemoryLabel{-18em}{0.75em}{0}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
%  \MemoryLabel{-10em}{0.75em}{0}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 12} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{5}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(5\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 4} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
%    \memory{4}{\unsealc}
    \memory{1}{\unsealc}
    \memory{1}{\retptrc}
    \memory{2}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{5}
    \MemoryLabel{-14em}{0.75em}{2}
    \MemoryLabel{-10em}{0.75em}{5}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  }
  \\
  {\tt 21} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-10em}{0.75em}{5}
 % \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_3\)}
  \MemoryLabel{-10em}{0.75em}{\(v_4\)}
  \MemoryLabel{-6em}{0.75em}{\(v_5\)}
  \\ &
  {\bf 22}
  ~$\cdots$
  &
  {\bf 24}
  ~$\cdots$
  \\
\end{tabular}
\end{center}

@@

\paragraph*{Control Flow Attacks}
\begin{figure}
  \centering
  \begin{subfigure}{.4\textwidth}
\begin{verbatim}
int *stash = 0;

int main() {
0:     int x;
1:     x = 1;
2[C]:  f();
3:     x = -x;
4[C]:  f();
5[R]:  return x;
}

void f() {
10:     if(!stash) {
        // store return pointer
11:       stash = $SP-1;
12:     } else {
13:       *($SP-1) = stash;
14:       stash = 0;
        }
15[R]:    return;
}
\end{verbatim}
  \end{subfigure}
  \begin{subfigure}{.5\textwidth}
    \begin{tabular}{l l l}
      {\tt 2} &
      \memoryaddrs{8em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1} \\
      {\tt 15} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\unsealc}[RetPtr]%
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1}
      \MemoryLabel{-15em}{0.75em}{\#3}
      \vspace{.5em} &
      \memory[3em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#3}
      \\
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{4}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{1} \\
      {\tt 10} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{3}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1}
      \MemoryLabel{-15em}{0.75em}{\#5} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
      \\
      {\tt 15} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1}
      \MemoryLabel{-15em}{0.75em}{\#3} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
      \\
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{-1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-19em}{0.75em}{\bf 1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#5} \\
    \end{tabular}
  \end{subfigure}
  \caption{A Control Flow Attack.}
  \label{fig:controlflow}
\end{figure}

%From the classic stack smashing attack to return oriented programming,
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow follows a typical structure
of matching sets calls and returns nested within on another.
Rather, they express what it means to protect data even in the presence of control-flow attacks.
Consider example \ref{fig:controlflow}.
Here we show return addresses explicitly as line numbers, in the form $\#n$.
Function {\tt f} is called twice, at lines 2 and 4; it should therefore return to lines
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to line 3 instead of 5; then it ends
up being called a third time, at which point it returns to line 5.

Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that we return
-1. But because of {\tt f}'s interference, we negate {\tt x} twice, and return 1.
This has the appearance of an integrity violation.  How do we capture this in our
integrity property?

The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs
to do more than just execute a return-labeled instruction: it must actually reach the
appropriate {\em return target} in {\tt main}. A return target is a predicate on states,
in this case one that holds on any state whose program counter points to the correct line.
We introduce a new piece of context: we keep a stack of return targets, and a callee is
only considered to have returned when it reaches a state that matches the caller's target.
A return target is associated with a set of sealed locations that will become unsealed
when it is reached, here represented by sharing a color with the sealed data.

In the example, on its first return, {\tt f} reaches its return target, because the
\(\PCname\) is 3 after its return. But on the second, {\tt f} does not properly return,
but rather changes its stack pointer (shown in light red) and jumps into
{\tt main}'s code. Then {\tt f}
attempts to overwrite the (still sealed) value of {\tt x}, violating integrity.

>>>>

int g() {
  int z;
  if (z == 5)
    return 10;
  else
    return 20;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.7\textwidth}
  {\small
\begin{verbatim}
1 main:  add $1,%sp      ; allocate frame
2        call f
3        mov %ra,-1(%sp) ; store x
4        call g
5        sub $1,%sp      ; deallocate frame
6        ret

10 f:    add $1,%sp      ; allocate frame
11       mov $5,-1(%sp)  ; initialize y
12       mov -1(%sp),%ra ; set return value
13       sub $1,%sp      ; deallocate frame
14       ret

20 g:    add $1,%sp      ; allocate frame
21       cmp $5,-1(%sp)  ; compare 
22       brneq #25       ; branch if not equal
23       mov $10,%ra     ; set return value
24       jmp #26
25       mov $20,%ra     ; set return value
26       sub $1,%sp      ; deallocate frame
27       ret 
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.65\textwidth}
\begin{center}
\begin{tabular}{l r | l}
  {\tt 2} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
%    \MemoryLabel{-18em}{0.75em}{0}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
%  \MemoryLabel{-10em}{0.75em}{0}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(v_1\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 12} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
 % \MemoryLabel{-19em}{0.75em}{0}
  \MemoryLabel{-10em}{0.75em}{5}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-10em}{0.75em}{\(5\)}
  \MemoryLabel{-6em}{0.75em}{\(v_2\)}
  \\
  {\tt 4} &
  \multicolumn{2}{c}{
    \memoryaddrs{8em}
%    \memory{4}{\unsealc}
    \memory{1}{\unsealc}
    \memory{1}{\retptrc}
    \memory{2}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-18em}{0.75em}{5}
    \MemoryLabel{-14em}{0.75em}{2}
    \MemoryLabel{-10em}{0.75em}{5}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  }
  \\
  {\tt 22} &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-10em}{0.75em}{5}
 % \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs{16em}
  \memory{1}{\unsealc}
  \memory{1}{\retptrc}
  \memory{2}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_3\)}
  \MemoryLabel{-10em}{0.75em}{\(v_4\)}
  \MemoryLabel{-6em}{0.75em}{\(v_5\)}
  \\ &
  {\bf 23}
  ~$\cdots$
  &
  {\bf 25}
  ~$\cdots$
  \\
\end{tabular}
\end{center}

@@

\paragraph*{Control Flow Attacks}
\begin{figure}
  \centering
  \begin{subfigure}{.3\textwidth}
{\small
\begin{verbatim}
int *stash = 0;

int main() {
  int x = 1;
  f();
  x = -x;
  f();
  return x;
}

void f() {
  int y;
  if (!stash) {
    stash = *(&y-1);
  } else {
    *(&y-1) = stash;
    stash = 0;
  }
  return;
}
\end{verbatim}
}
\end{subfigure}
  \begin{subfigure}{.65\textwidth}
{\small
\begin{verbatim}
   stash: .word 0
1  main:  add $1,%sp        ; allocate frame
2         mov $1,-1(%sp)    ; initialize x
3         call f
4         neg -1(%sp)       ; x = -x 
5         call f           
6         mov -1(%sp),%ra   ; set return value
7         sub $1,%sp        ; deallocate frame
6         ret

10 f:     add $1,%sp        ; allocate frame
11        cmp stash, $0     ; compare
12        breq #15          ; branch if equal
13        mov -2(%sp),stash ; update stash
14        jmp #17
15        mov stash,-2(%sp) ; retrieve stash
16        mov $0,stash
17        sub $1,%sp        ; deallocate frame
18        ret 
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}{.4\textwidth}
    \begin{tabular}{l l l}
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\unsealc}[RetPtr]%
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1}
      \MemoryLabel{-11em}{0.75em}{\#4}
      \vspace{.5em} &
      \memory[3em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#4}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 11} &
      \memoryaddrs{16em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#6} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
      \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#4} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6}
      \\
      {\tt 5} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{\bf 1} &
      \memory[3em]{1}{\mainsealc}
      \MemoryLabel{-3em}{0.75em}{\(\PCname\) = \#6} \\
    \end{tabular}
  \end{subfigure}
  \caption{A Control Flow Attack.}
  \label{fig:controlflow}
\end{figure}

%From the classic stack smashing attack to return oriented programming,
One of the most pernicious ways to attack the stack is by hijacking control flow.
Our stack safety properties do not assume that control flow follows a typical structure
of matching sets calls and returns nested within on another.
Rather, they express what it means to protect data even in the presence of control-flow attacks.
Consider example \ref{fig:controlflow}.
Here we show return addresses explicitly as line numbers, in the form $\#n$.
Function {\tt f} is called twice, at instructions 2 and 4; it should therefore return to instructions
3 and 5, respectively. But the first time, before returning, it stashes its
return pointer away, and the second time it returns to instruction 3 instead of 5; then it ends
up being called a third time, at which point it finally returns to instruction 5.

Reasoning about {\tt main}, we expect that {\tt x} is negated once, so that {\tt main} returns
-1. But because of {\tt f}'s interference, {\tt x} is negated twice, and {\tt main} returns 1.
This has the appearance of an integrity violation.  How do we capture this in our
integrity property?

The key intuition is that to be considered as having returned to {\tt main}, {\tt f} needs
to do more than just execute a return-labeled instruction: it must actually reach the
appropriate {\em return target} in {\tt main}. A return target is a predicate on states,
in this case one that holds on any state whose program counter points to the correct instruction.
We introduce a new piece of context: we keep a stack of return targets, and a callee is
only considered to have returned when it reaches a state that matches the caller's target.
A return target is associated with a set of sealed locations that will become unsealed
when it is reached, here represented by sharing a color with the sealed data.

In the example, on its first return, {\tt f} reaches its return target, because the
\(\PCname\) is 3 after its return. But {\tt f}'s second return is to the wrong instruction
(3 rather than 5) because the stack pointer has been overwritten (shown in light red).
Thus {\tt x} is still sealed and {\tt f}'s attempt to overwrite it at instruction 4
violates the integrity property.
